Conversations in time: interactive visualisation to explore structured temporal data

### 2020-08-26 tsibbletalk.Rmd
added narratives to code chunks

This dataset contains a three-level hierarchy: the root node is implicitly Australia, and geographically disaggregated to states and lower-level tourism regions. A new handy function `plotly_key_tree()` has been implemented to address the need of hierarchical discovery arising from the data. It visually interprets hierarchies in the shared tsibble's `spec` as a tree view, made with \CRANpkg{plotly}. The following code line produces the linked tree diagram and fills the left panel of Figure \ref{fig:tourism-linking-fig}.

@@

The visual of tree hierarchy untangles a group of related series and snapshots the data organisation from a bird's eye view. It is accompanied with other two plots to compose an ensemble of coordinated graphics for multi-facetted exploration, shown as Figure \ref{fig:tourism-linking-fig}. The time series and scatter plots of Figure \ref{fig:tourism-linking-fig} are the interactive realisation of Figure \ref{fig:highlight-retail}. Since all plots are derived from one shared tsibble data source, they are self-linking multiple views. For example, selecting the leaf node "Melbourne" fires up four respective series of the `Purpose` category in other subplots, characterising "categorical linking". Press the "Shift" key to switch on persistent selection and simultaneously select the parent node "Western Australia" to trigger its children. The domestic tourism performance in Melbourne can be quickly compared against the whole state of Western Australia. It makes navigation simpler than conditional panels, and also enables persistent selection for comparisons across groups.

@@

```{r tourism-linking, eval = FALSE}
```

>>>>

This dataset contains a three-level hierarchy: the root node is implicitly Australia, and geographically disaggregated to states and lower-level tourism regions. A new handy function `plotly_key_tree()` has been implemented to address the need of hierarchical discovery arising from the data. It interprets hierarchies in the shared tsibble's `spec` as a tree view, built with \CRANpkg{plotly}. The following code line produces the linked tree diagram and fills the left panel of Figure \ref{fig:tourism-linking-fig}. The visual of tree hierarchy untangles a group of related series and snapshots the data organisation from a bird's eye view.

@@

The tree plot provides backbones of the data, and much flesh yet to be attached. Small multiples of time series lines are composed and placed at the top right of Figure \ref{fig:tourism-linking-fig} to unpack the temporal trend across regions by purposes of trips. The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code.

```{r tourism-series, echo = TRUE, eval = FALSE}
```

To tease apart these overlaid time series, they are funnelled through the `features()` S3 method to extract some key characteristics, including the measurements of trend and seasonality. A scatterplot is populated from these statistics for each series.

```{r tourism-scatter, echo = TRUE, eval = FALSE}
```

Lastly, three graphics are composed as an ensemble of coordinated views for multi-facetted exploration, shown as Figure \ref{fig:tourism-linking-fig} (the interactive realisation of Figure \ref{fig:highlight-retail}). Routine functions bring about new interaction with temporal data on the client side (i.e. without Shiny).

```{r tourism-multi, echo = TRUE, eval = FALSE}
```

Since all plots are stemmed from one shared tsibble data source, they are self-linking views. Nodes, lines, and points are clickable. For example, selecting the leaf node "Melbourne" fires up four respective series of the `Purpose` category in other subplots, characterised by "categorical linking". Press the "Shift" key to switch on persistent selection, and simultaneously select the parent node "Western Australia" to trigger its children. The domestic tourism performance in Melbourne can be quickly compared against the whole state of Western Australia. It makes navigation simpler than conditional panels, and also enables persistent selection for comparisons across groups.

@@

### 2020-08-26 tsibbletalk.tex
added narratives to code chunks

This dataset contains a three-level hierarchy: the root node is
implicitly Australia, and geographically disaggregated to states and
lower-level tourism regions. A new handy function
\texttt{plotly\_key\_tree()} has been implemented to address the need of
hierarchical discovery arising from the data. It visually interprets
hierarchies in the shared tsibble's \texttt{spec} as a tree view, made
with \CRANpkg{plotly}. The following code line produces the linked tree
diagram and fills the left panel of Figure
\ref{fig:tourism-linking-fig}.

@@

The visual of tree hierarchy untangles a group of related series and
snapshots the data organisation from a bird's eye view. It is
accompanied with other two plots to compose an ensemble of coordinated
graphics for multi-facetted exploration, shown as Figure
\ref{fig:tourism-linking-fig}. The time series and scatter plots of
Figure \ref{fig:tourism-linking-fig} are the interactive realisation of
Figure \ref{fig:highlight-retail}. Since all plots are derived from one
shared tsibble data source, they are self-linking multiple views. For
example, selecting the leaf node ``Melbourne'' fires up four respective
series of the \texttt{Purpose} category in other subplots,
characterising ``categorical linking''. Press the ``Shift'' key to
switch on persistent selection and simultaneously select the parent node
``Western Australia'' to trigger its children. The domestic tourism
performance in Melbourne can be quickly compared against the whole state
of Western Australia. It makes navigation simpler than conditional
panels, and also enables persistent selection for comparisons across
groups.

>>>>

This dataset contains a three-level hierarchy: the root node is
implicitly Australia, and geographically disaggregated to states and
lower-level tourism regions. A new handy function
\texttt{plotly\_key\_tree()} has been implemented to address the need of
hierarchical discovery arising from the data. It interprets hierarchies
in the shared tsibble's \texttt{spec} as a tree view, built with
\CRANpkg{plotly}. The following code line produces the linked tree
diagram and fills the left panel of Figure
\ref{fig:tourism-linking-fig}. The visual of tree hierarchy untangles a
group of related series and snapshots the data organisation from a
bird's eye view.

@@

The tree plot provides backbones of the data, and much flesh yet to be
attached. Small multiples of time series lines are composed and placed
at the top right of Figure \ref{fig:tourism-linking-fig} to unpack the
temporal trend across regions by purposes of trips. The shared tsibble
data can be directly piped into \CRANpkg{ggplot2} code.

\begin{Schunk}
\begin{Sinput}
p_tr <- tourism_shared %>%
  ggplot(aes(x = Month, y = Trips)) +
  geom_line(aes(group = Region), alpha = .5, size = .4) +
  facet_wrap(~ Purpose, scales = "free_y") +
  scale_x_yearmonth(date_breaks = "5 years", date_labels = "%Y")
\end{Sinput}
\end{Schunk}

To tease apart these overlaid time series, they are funnelled through
the \texttt{features()} S3 method to extract some key characteristics,
including the measurements of trend and seasonality. A scatterplot is
populated from these statistics for each series.

\begin{Schunk}
\begin{Sinput}
tourism_feat <- tourism_shared %>%
  features(Trips, feat_stl)
p_br <- tourism_feat %>%
  ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
  geom_point(aes(group = Region))
\end{Sinput}
\end{Schunk}

Lastly, three graphics are composed as an ensemble of coordinated views
for multi-facetted exploration, shown as Figure
\ref{fig:tourism-linking-fig} (the interactive realisation of Figure
\ref{fig:highlight-retail}). Routine functions bring about new
interaction with temporal data on the client side (i.e.~without Shiny).

\begin{Schunk}
\begin{Sinput}
subplot(p_l,
  subplot(
    ggplotly(p_tr, tooltip = "Region", width = 1100),
    ggplotly(p_br, tooltip = "Region", width = 1100),
    nrows = 2),
  widths = c(.4, .6)) %>%
  highlight(dynamic = TRUE)
\end{Sinput}
\end{Schunk}

Since all plots are stemmed from one shared tsibble data source, they
are self-linking views. Nodes, lines, and points are clickable. For
example, selecting the leaf node ``Melbourne'' fires up four respective
series of the \texttt{Purpose} category in other subplots, characterised
by ``categorical linking''. Press the ``Shift'' key to switch on
persistent selection, and simultaneously select the parent node
``Western Australia'' to trigger its children. The domestic tourism
performance in Melbourne can be quickly compared against the whole state
of Western Australia. It makes navigation simpler than conditional
panels, and also enables persistent selection for comparisons across
groups.
